I Compensate, therefore I Am \\ (accurate for DNN verification)

### 2024-01-27 Paper/proofs.tex
proof of th2

@@

We assume the fairly light hypothesis (H1) that the DNN is such that for every 2 paths $\rho_1,\rho_2$ (not necessarily disjoint) from any $m$ to any $n$ at least 2 layers apart, there exists another path $\rho_3$ from $m$ to $n$ (with non zero-weight) totally disjoint from $\rho_1$ and $\rho_2$. Further we assume without loss of generality that every node can be reached from at least one input with a path of non-zero weight (otherwise we just remove this node altogether).
We also assume without loss of generality that there is no node which can only reach $z$ with path of weight $0$, as such nodes have no contribution to the value $\val(z)$. We denote $m<n$ to mean that there exists a path with non zero weight from $m$ to $n$.

@@

To prove Theorem \ref{th2}, we can then optimize $Y$ first on its own (considering the sign of every node): use the same argument in Section A, the proof of Theorem \ref{th1}, we can maximize all $n$ such that $S(n)=1$ and minimize all $n$ such that $S(n)=-1$ at the same time, i.e., with the same input values. And for every node $n$ in $Y$, the value computed is the same as actual lower and upper bounds of $n$.

>>>>

We now prove Theorem \ref{th2} under the fairly light {\em well-connected hypothesis} (H1) that the DNN is such that for every 2 (not necessarily disjoint) paths $\rho_1,\rho_2$ with non-zero weight with the same source $m$ and the same target $n$ at least 2 layers apart, there exists another path $\rho_3$ from $m$ to $n$ (with non zero-weight) totally disjoint from $\rho_1$ and $\rho_2$ (except at the source and target). This hypothesis is used to remove corner cases which will not happen in actual DNNs, which all are well-connected.

Further we assume without loss of generality that every node can be reached from at least one input with a path of non-zero weight (otherwise we just remove this node altogether).
We also assume without loss of generality that there is no node which can only reach $z$ with path of weight $0$, as such nodes have no contribution to the value $\val(z)$. We denote $m<n$ to mean that there exists a path with non zero weight from $m$ to $n$.

Under (H1), we have the following Lemma:

\begin{lemma}
  Let $n$ be on a {\em weakly} compensating pair $(\rho_1,\rho_2)$ of paths, 
  except an input or $z$. Then there exists a compensating pair of paths 
 $(\rho,\rho')$ including $n$, not as its source or its target.
\end{lemma}

\begin{proof}
Let $(s,t)$ be the source and the target of $(\rho_1,\rho_2)$.
Let an input $a$ such that $a<s$ with path $\rho_0$ between $a$ and $s$,
and $\rho_3$  from $t$ to $z$.
Applying (H1) to the pair 
$(\rho_0 \cdot \rho_1 \cdot \rho_3, \rho_0\cdot \rho_2 \cdot \rho_3)$, 
we have another (non zero weight) path $\rho'$ from $a$ to $z$ totally disjoint with these two path.
The weights of $(\rho_0 \cdot \rho_1 \cdot \rho_3, \rho_0\cdot \rho_2 \cdot \rho_3)$
are of opposite sign because $(\rho_1,\rho_2)$ is weakly compensating, so one of them must be of opposite sign than the weight of $\rho'$. Say its $\rho = \rho_0 \cdot \rho_1 \cdot \rho_3$
else we pick $\rho=\rho_0\cdot \rho_2 \cdot \rho_3$.
Then $(\rho,\rho')$ is a pair of compensating path from $a$ to $z$, and as $n$ is neither an input nor $z$, it is strictly on $\rho$.
\end{proof}

It suffices to invoke Theorem \ref{thm:2} in order to get the proof for Theorem \ref{th2},
as the set $X$ considered in both are actually the same.

\end{document}

@@

We assume the fairly light hypothesis (H1) that the DNN is such that for every 2 paths $\rho_1,\rho_2$ (not necessarily disjoint) from any $m$ to any $n$ at least 2 layers apart, there exists another path $\rho_3$ from $m$ to $n$ (with non zero-weight) totally disjoint from $\rho_1$ and $\rho_2$.

Further we assume without loss of generality that every node can be reached from at least one input with a path of non-zero weight (otherwise we just remove this node altogether).
We also assume without loss of generality that there is no node which can only reach $z$ with path of weight $0$, as such nodes have no contribution to the value $\val(z)$. We denote $m<n$ to mean that there exists a path with non zero weight from $m$ to $n$.

@@

To prove Theorem \ref{th2}, we can then optimize $Y$ first on its own (considering the sign of every node): we first maximize all $n$ such that $S(n)=1$ and minimize all $n$ such that 
$S(n)=-1$ at the same time, i.e., with the same input values. And for every node $n$ in $Y$, the value computed is the same as actual lower and upper bounds of $n$.
